Relationships in the extremes and their influence on competition and coexistence

### 2021-09-20 paper/Paper.Rnw
desert annual as example

%1) Topic sentence: To make the above ideas more concrete, we review a relevant example from the literature that may be
%influenced as described above. 

@@

To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and, following our simulated example (Figure \ref{fig:pedagog}), the species persist in coexistence. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.

>>>>

%1) Topic sentence: 
To make the above ideas more concrete, we review a relevant example from the literature that may be
influenced as previously described. 
We introduce a simplified example of an Arizona desert annual plant community from \cite{Chesson_2012}. 

@@

The community follows the criteria required for temporal storage effects to operate. Different species dominate and have high germination rates depending on rainfall timing. For example, Baileya multiradiata dominates if sufficient rainfall is reached in the early germination season. Eriogonum abertianum dominates is sufficient rainfall is reached in the late germination season. Competition is influenced by resource consumption of seed-producing adult plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the desert annual community also fluctuate year to year. The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the desert annual response relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the desert annual environmental response relationship may become more correlated in the lower values or left tail associated.

To make the above ideas more concrete, we review a relevant example from the literature that may be
influenced as previously described. 
We introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
%2) Make it concrete - name the system, name the organisms, name the seasons. 
To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and, following our simulated example (Figure \ref{fig:pedagog}), the species persist in coexistence. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.
